\section*{Problem 2}

To prove that MOST NON-NEGATIVE SOLUTION is in NP, we need to provide the recipe for a polynomially-sized certificate which can be verified in polynomial time. The certificate is just a set $K$ of $k$ variables which will take non-negative values. Note that the size of the problem in bits is $\text{poly}(n \log U)$ where $U$ is the largest integer (in absolute value) in $A$ or $b$. To verify the certificate we need to verify that there is some $x$ where 

\begin{align*}
Ax &= b\\
x_i &\geq b \quad \forall i \in K
\end{align*}

which can be done in polynomial time of the input size using the ellipsoid algorithm.\\

To prove that MOST NON-NEGATIVE SOLUTION is NP-Complete, we first rephrase the problem be: Given a positive integer $k$ and a system of linear equations $Ax = b$, is there a solution that at most $m-k$ variables are negative. Call the rephrased problem MODIFIED MOST NON-NEGATIVE SOLUTION. We now show a 3SAT problem with $n$ variables and $m$ clauses can be reduced to MODIFIED MOST NON-NEGATIVE SOLUTION which allows at most $n$ negative variables in polynomial time.\\

We construct four variables $x_i$, $\bar{x_i}$, $s_i$ and $\bar{s_i}$ for each variable in 3SAT and one variable $t_j$ for each clause $j$ in 3SAT. We add the following constraints to $A$ and $b$. 
\begin{align}
x_i + \bar{x_i} &= 1 \hspace{5mm} \forall i\\
x_i - s_i &= 1 \hspace{5mm} \forall i\\
\bar{x_i} - \bar{s_i} &= 1\hspace{5mm} \forall i\\
x_a + x_b + x_c - t_j &= 1 \hspace{5mm} \forall j = (x_a\lor x_b \lor x_c)
\end{align}
We are proving that a solution to MODIFIED MOST NON-NEGATIVE SOLUTION problem (MMNNS) returns a satisfying assignment to 3SAT.
\begin{lemma} At least one of $s_i$ and $\bar{s_i}$ is negative in any solution to the MMNNS problem. \end{lemma}
\begin{proof} If both $s_i$ and $\bar{s_i}$ are non-negative, then the sum of eq (2) and (3) will violate eq (1). Therefore, at least one of $s_i$ and $\bar{s_i}$ must be negative.\end{proof}
\begin{lemma} At most one of $s_i$ and $\bar{s_i}$ is negative.\end{lemma}
\begin{proof} Any solution to MMNNS problem has at most $n$ negative variables. From \textbf{Lemma 1}, we showed that at least one of $s_i$ and $\bar{s_i}$ is negative; there are $n$ pairs of them. Therefore, at most one of $s_i$ and $\bar{s_i}$ is negative. \end{proof}
\begin{corollary} All variables except $s_i$ and $\bar{s_i}$ must be non-negative for any solution to MODIFIED MOST NON-NEGATIVE SOLUTION problem.
\end{corollary}
\begin{lemma}Exactly one of $x_i$ and $\bar{x_i}$ is 1 and the other is 0 in any solution to MMNNS.\end{lemma}
\begin{proof} From \textbf{Lemma 1},  \textbf{Lemma 2} and eq (2) and (3), we know that one of $x_i$ and $\bar{x_i}$ is strictly greater than 1 and the other one is between 0 and 1. From eq (1), we conclude that one of $x_i$ and $\bar{x_i}$ is 1 and the other is 0.\end{proof}
\textbf{Backward Direction:} A solution to MODIFIED MOST NON-NEGATIVE SOLUTION can be decoded into a satisfying truth assignment of 3SAT.
\begin{proof}
By \textbf{Lemma 3}, we know for any feasible candidate of MODIFIED MOST NON-NEGATIVE SOLUTION, exactly one of $x_i$ and $\bar{x_i}$ is 1. We can assign variable $x_i$  in 3SAT to be true if the corresponding variable $x_i$ in MODIFIED MOST NON-NEGATIVE SOLUTION is 1 or false if $\bar{x_i}$ is 1.\\
This is a satisfying truth assignment for 3SAT because: (1) each variable is assigned either truth or false but not both; (2) for each clause $c_j = (x_a \lor x_b \lor x_c)$, at least one of the literals has to have truth assignment, otherwise $t_j$ will be forced to be negative, which contradicts \textbf{Corollary 1}.  
\end{proof}
\textbf{Forward Direction:} A satisfying truth assignment of 3SAT can be transformed into a feasible candidate of MODIFIED MOST NON-NEGATIVE SOLUTION.
\begin{proof} For variable $x_i$ in 3SAT whose value is true, assign $x_i$ in MODIFIED MOST NON-NEGATIVE SOLUTION to be 1 and $\bar{x_i}$ to be 0, otherwise assignment $\bar{x_i}$ to be 1 and $x_i$ to be 0. Correspondingly, we have $s_i = x_i - 1$ and $\bar{s_i} = \bar{x_i} -1$. The value of $t_j$ is determined by number of truth variable in the clause $t_j=\{0,1,2\}$ $\forall j$. \\
This is a solution to MODIFIED MOST NON-NEGATIVE SOLUTION because all variables are non-negative except $s_i$ and $\bar{s_i}$ and in fact exactly one of the two is negative by construction. Therefore, there are exactly $n$ variables that are negative for this candidate.  
\end{proof}